---
title: "Dataopgave"
author: "Tiemo Thiess"
date: "22/6/2020"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I started the task by generally exploring the data and thinking about how this data can help to solve a business problem for PFA. A common EDA does not seem to be appropriate as the data is already aggregated and presented as analysis results. Because of this, I choose to do a correlation analysis of variables that may be of interest for PFA such as the relationship between offering treatment programs and the overall workability. Ideally I would like to use the raw data of the 50000 respondents, but as this is not available, I am basing the analysis on the aggregated data. In the end of the analysis, I am using the results of the correlation analysis to create clusters of job profiles that could be a potential target group for PFA products and services. 

## Load packages and data
Instead of joining the received datasets with the ones from prior years, I used the full dataset that was accesible via the NFA.
```{r, message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(cluster)  
library(corrplot)
library(ggplot2)

data_raw<- read_excel("~/Documents/na18_tableau_resultater_Migrated Data.xls")
```
## Filter data
I searched trough the different topics and found a subset of topics that could be potentially interesting for PFA. Also, I focus on the main group of job profiles. 
```{r}
topics_list <- c("Arbejdsevne",
            "Arbejdsrelateret sygdom",
            "Begrænset i arbejdet pga. smerter" ,
            "Depressive symptomer",
            "Mental sundhed",
            "Selvvurderet helbred",
            "Tilbud om behandlingsordning",
            "Tilbud om motionsfaciliteter eller ugentlig motion",
            "Tilbud om rygestop",
            "Tilbud om små motionsaktiviteter i dagligdagen",
            "Tilbud om sund kost/kostvejledning",
            "Tilbud om sundhedstjek",
            "Vitalitet")

names(data_raw) <- gsub(" ", "_", names(data_raw)) # remove spaces
names(data_raw) <- tolower(names(data_raw)) # tranform to lower case column names
data <- data_raw %>% 
  filter(main_group == 'Job') %>%
  filter(title %in% topics_list)
```

## Tidy data
To perform a correlation analysis, I reshaped the data from a long into a tidy wide format in which each row is a different job profile and each column is a variable that is related to the survey questions.
```{r}
# From long to wide
data_wide <- data %>%
  select(group, question, score) %>%
  drop_na() %>%
  mutate(id=1:n()) %>%
  spread(question, score) %>%
  select(-id) %>%
  group_by(group) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

head(data_wide)
```
## Correlation analysis
```{r}
scaled <- scale(data_wide[, -1])
cor <- cor(scaled)
corrplot(cor, method="color", tl.col = "black", addCoef.col = "black", number.cex = 0.5, diag=FALSE)
```

## Insights 
From the correlation analysis, I derived some insights.

1. Offering sports or general treatment opportunities seems to have positive effects on workability, work related sickness, depression, pain, and vitality

2. Offering support for stopping smoking has only moderate effects on most variables but positive effects on the self-assessed health

3. Offering general health checks seems to have less benefitial effects than general treatment programs and sports offerings


```{r}
# Development over the years
data_wide_years <- data %>%
  select(question, year, score) %>%
  mutate(year = as.character(year)) %>%
  drop_na() %>%
  mutate(id=1:n()) %>%
  spread(question, score) %>%
  select(-id) %>%
  group_by(year) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  mutate_if(is.numeric, scale) %>%
  mutate(year = as.numeric(year))

summary(data_wide_years)
head(data_wide_years)

# Visualization
ggplot(data_wide_years, aes(x=year), group = 1) + 
  geom_line(aes(y = ArbEvne, colour = "ArbEvne")) + 
  geom_line(aes(y = Behandling, colour = "Behandling")) +
  geom_line(aes(y = Motionsakt, colour = "Montionsakt")) + 
  geom_line(aes(y = SelvHelbred, colour = "SelvHelbred")) +
  geom_line(aes(y = Rygestop, colour = "Rygestop"))
```


```{r} 
# Cluster Analysis
scaled_filt <- dplyr::select(as.data.frame(scaled), SelvHelbred, Rygestop)
k <- kmeans(scaled_filt[, ], centers = 2, nstart = 25)
k

# Elbow Plot
wssplot <- function(scaled_filt, nc=15, seed=1234){
  wss <- (nrow(scaled_filt)-1)*sum(apply(scaled_filt,2,var))
  for (i in 2:nc){
    set.seed(seed)
    wss[i] <- sum(kmeans(scaled_filt, centers=i)$withinss)}
  plot(1:nc, wss, type="b", xlab="Number of Clusters",
       ylab="Within groups sum of squares")}

wssplot(scaled_filt, nc=10) 

k2 <- kmeans(scaled_filt[, ], centers = 5, nstart = 25)
k2 #Insight 2 jobs and industries in cluster 1 are potential target groups

# Target group
clust_data <- cbind(data_wide[, 1], k2$cluster) %>%
  filter(k2$cluster == 2)
names(clust_data) <- c("group", "cluster")
clust_data
```

